---
title: Translating GEMMA’s Newton-Raphson Algorithm from C++ to R
author: Frederick Boehm
date: '`r lubridate::now()`'
slug: translating-gemma-s-newton-raphson-algorithm-from-c-to-r
draft: 'true'
categories:
  - Computing
tags:
  - Newton-Raphson
  - Linear mixed effects models
  - Variance components
  - R
---

## Overview

We want to translate GEMMA's C++ code in the function `MphNR` (in the file `mvlmm.cpp`) into R code as a way to better understand the calculations and the algorithm.

### `MphNR` C++ Code

`MphNR` is defined in [these lines](https://github.com/xiangzhou/GEMMA/blob/3ed15d7061fdac57c07da0fba087e9e69ac0ec42/src/mvlmm.cpp#L2504-L2662) in the GEMMA Github repository. 

### R code for `MphNR`

```{r}
#double MphNR (const char func_name, const size_t max_iter, const double max_prec, const gsl_vector *eval, const gsl_matrix *X, const gsl_matrix *Y, gsl_matrix *Hi_all, gsl_matrix *xHi_all, gsl_matrix *Hiy_all, gsl_matrix *V_g, gsl_matrix *V_e, gsl_matrix *Hessian_inv, double &crt_a, double &crt_b, double &crt_c)
#{
MphNR <- function(func_name = "R", max_iter, max_prec, eval, X, Y, Hi_all, xHi_all, Hiy_all, V_g, V_e){
  n_size <- length(eval)
  c_size <- nrow(X)
  d_size <- nrow(Y)
  dc_size <- d_size * c_size
  v_size <- d_size * (d_size + 1) / 2
  XXt <- X %*% t(X)
  log(det(XXt)) -> lndetXXt
  if (func_name == "R"){
    logl_const <- - 0.5 * (n_size - c_size) * d_size * log(2 * pi) + 0.5 * d_size * lndetXXt
  }
  for (t in 1:max_iter){
    Vg_save <- V_g
    Ve_save <- V_e
    step_scale <- 1
    step_iter <- 0
    while((flag_pd == 0 | logl_new<logl_old | logl_new - logl_old > 10 ) & step_iter < 10){
      V_g <- Vg_save
      V_e <- Ve_save
      if (t > 1){UpdateVeVg()}
      flag_pd <- matrixcalc::is.positive.definite(V_g) & matrixcalc::is.positive.definite(V_e)
			if (flag_pd == 1) {
			  if (func_name=='R') {
					logl_new <- logl_const - 0.5 * logdet_H - 0.5 * logdet_Q - 0.5 * yPy
			  }
			}
      step_scale <- step_scale / 2
      step_iter <- step_iter + 1
    } # end of while loop
    if (t > 1){
      if (logl_new < logl_old | flag_pd == 0) {
				V_g <- Vg_save
				V_e <- Ve_save
				break
      }
      if (logl_new - logl_old < max_prec) {
				break
      }
    logl_old <- logl_new
    foo <- CalcDev(func_name = "R", eval, Qi, Hi, xHi, Hiy, QixHiy) ## FIX HERE
    foo[[1]] -> Hessian_inv
    foo[[2]] -> gradient
    }
  } # end of loop over t
#   //mutiply Hessian_inv with -1.0
# 	//now Hessian_inv is the variance matrix
# 	gsl_matrix_scale (Hessian_inv, -1.0);
  
  return(logl_new)
} # end of function
```

The [do-while loop](https://github.com/xiangzhou/GEMMA/blob/3ed15d7061fdac57c07da0fba087e9e69ac0ec42/src/mvlmm.cpp#L2553-L2598) is a construction that I haven't seen before in C++. We can translate it into R by using a `while` loop.

`flag_pd` is an indicator for positive-definite matrices.

### `CalcHiQi`

`CalcHiQi` is defined [here](https://github.com/xiangzhou/GEMMA/blob/898770748aa87ad89a5ca10282141fae7f4cbbd5/src/mvlmm.cpp#L961-L1022) in GEMMA.

```{r}
CalcHiQi <- function(eval, X, V_g, V_e){
  n_size <- length(eval)
  c_size <- nrow(X)
  d_size <- nrow(V_g)
  ep_out <- eigen_proc(V_g, V_e) # I defined eigen_proc() in my earlier blog post from may 24, 2017
  ep_out[[1]] -> logdet_Ve
  ep_out[[2]] -> UltVeh
  ep_out[[3]] -> UltVehi
  ep_out[[4]] -> D_l
  logdet_H <- n_size * logdet_Ve
  for (k in 1:n_size){
    #delta=gsl_vector_get (eval, k);
    delta <- eval[k]
    #gsl_matrix_memcpy (mat_dd, UltVehi);
    mat_dd <- UltVehi
    #for(size_t i=0; i<d_size; i++) {
		for (i in 1:d_size){
    	#dl=gsl_vector_get(D_l, i);
      dl <- D_l[i]
    	#d=delta*dl+1.0;
      d <- delta * dl + 1
			#gsl_vector_view mat_row=gsl_matrix_row (mat_dd, i);
			mat_row <- mat_dd[i, ]
      #gsl_vector_scale (&mat_row.vector, 1.0/d);
      mat_row <- mat_row / d
      mat_row -> mat_dd[i, ]
			#logdet_H+=log(d);
      logdet_H <- logdet_H + log(d)
		}
    #gsl_matrix_view Hi_k=gsl_matrix_submatrix(Hi_all, 0, k*d_size, d_size, d_size);
		Hi_k <- Hi_all[1:d_size, ((k - 1) * d_size + 1):(k * d_size)]
    #gsl_blas_dgemm(CblasTrans, CblasNoTrans, 1.0, UltVehi, mat_dd, 0.0, &Hi_k.matrix);
		Hi_k <- t(UltVehi) %*% mat_dd
		Hi_k -> Hi_all[1:d_size, ((k - 1) * d_size + 1):(k * d_size)]
  }
  #logdet_Q=CalcQi (eval, D_l, X, Qi)-(double)c_size*logdet_Ve;
  cq_out <- calc_qi(eval, D_l, X)
  cq_out[[1]] -> Qi
  cq_out[[2]] - c_size * logdet_Ve -> logdet_Q
  #for (size_t i=0; i<c_size; i++) {
	for (i in 1:c_size){
  	#for (size_t j=0; j<c_size; j++) {
		for (j in 1:c_size){
	  	#gsl_matrix_view Qi_sub=gsl_matrix_submatrix (Qi, i*d_size, j*d_size, d_size, d_size);
			Qi_sub <- Qi[((i - 1) * d_size + 1):(i * d_size), ((j - 1) * d_size + 1):(j * d_size)]
		  #if (j<i) {
			if (j < i){
				#gsl_matrix_view Qi_sym=gsl_matrix_submatrix (Qi, j*d_size, i*d_size, d_size, d_size);
				Qi_sym <- Qi[((j - 1) * d_size + 1):(j * d_size), ((i - 1) * d_size + 1):(i * d_size)]
			  #gsl_matrix_transpose_memcpy (&Qi_sub.matrix, &Qi_sym.matrix);
				t(Qi_sym) -> Qi_sub
				Qi_sub -> Qi[((i - 1) * d_size + 1):(i * d_size), ((j - 1) * d_size + 1):(j * d_size)]
			#} else {
			} else {
				#gsl_blas_dgemm(CblasNoTrans, CblasNoTrans, 1.0, &Qi_sub.matrix, UltVeh, 0.0, mat_dd);
				mat_dd <- Qi_sub %*% UltVeh
			  #gsl_blas_dgemm(CblasTrans, CblasNoTrans, 1.0, UltVeh, mat_dd, 0.0, &Qi_sub.matrix);
				t(UltVeh) %*% mat_dd -> Qi_sub
				Qi_sub -> Qi[((i - 1) * d_size + 1):(i * d_size), ((j - 1) * d_size + 1):(j * d_size)]
			}
		}
	}
  return(list(logdet_H, Qi))
}
  
```





### `Calc_Hiy_all`

```{r}
#void Calc_Hiy_all (const gsl_matrix *Y, const gsl_matrix *Hi_all, gsl_matrix *Hiy_all)
#{
Calc_Hiy_all <- function(Y, Hi_all){
#	gsl_matrix_set_zero (Hiy_all);
  

#	size_t n_size=Y->size2, d_size=Y->size1;
  ncol(Y) -> n_size
  nrow(Y) -> d_size
  Hiy_all <- matrix(data = 0, nrow = d_size, ncol = n_size)
#	for (size_t k=0; k<n_size; k++) {
  for (k in 1:n_size){
      #		gsl_matrix_const_view Hi_k=gsl_matrix_const_submatrix(Hi_all, 0, k*d_size, d_size, d_size);
    Hi_k <- Hi_all[1:d_size, ((k - 1) * d_size):(k * d_size)]
#		gsl_vector_const_view y_k=gsl_matrix_const_column(Y, k);
    y_k <- Y[, k]
#		gsl_vector_view Hiy_k=gsl_matrix_column(Hiy_all, k);
    Hiy_k <- Hiy_all[, k]
    
#		gsl_blas_dgemv (CblasNoTrans, 1.0, &Hi_k.matrix, &y_k.vector, 0.0, &Hiy_k.vector);
    Hiy_k <- Hi_k + y_k %*% t(y_k)
    Hi_k -> Hi_all[1:d_size, ((k - 1) * d_size):(k * d_size)]
    Hiy_k -> Hiy_all[, k]
#	}
  }
#	return;
#}
  return(Hiy_all)
}
```




### `Calc_xHi_all`

```{r}
#void Calc_xHi_all (const gsl_matrix *X, const gsl_matrix *Hi_all, gsl_matrix *xHi_all)
#{
Calc_xHi_all <- function(X, Hi_all){
  

	#gsl_matrix_set_zero (xHi_all);

	#size_t n_size=X->size2, c_size=X->size1, d_size=Hi_all->size1;
  n_size <- ncol(X)
  c_size <- nrow(X)
  d_size <- nrow(Hi_all)

#	double d;

#	for (size_t k=0; k<n_size; k++) {
  for (k in 1:n_size){
    
  	#	gsl_matrix_const_view Hi_k=gsl_matrix_const_submatrix(Hi_all, 0, k*d_size, d_size, d_size);
    Hi_all[1:d_size, ((k - 1) * d_size + 1):(k * d_size)] -> Hi_k

		#for (size_t i=0; i<c_size; i++) {
		for (i in 1:c_size){
		  
    	#d=gsl_matrix_get (X, i, k);
			d <- X[i, k]
		  #gsl_matrix_view xHi_sub=gsl_matrix_submatrix(xHi_all, i*d_size, k*d_size, d_size, d_size);
			xHi_sub <- xHi_all[((i - 1) * d_size + 1):(i * d_size), ((k - 1) * d_size + 1):(k * d_size)]
			#gsl_matrix_memcpy(&xHi_sub.matrix, &Hi_k.matrix);
			Hi_k -> xHi_sub
			#gsl_matrix_scale(&xHi_sub.matrix, d);
			xHi_sub <- xHi_sub / d
			xHi_sub -> xHi_all[((i - 1) * d_size + 1):(i * d_size), ((k - 1) * d_size + 1):(k * d_size)]
			
		}
    Hi_all[1:d_size, ((k - 1) * d_size + 1):(k * d_size)] <- Hi_k

	}

	#return;
  return(xHi_all)
}
```


### `Calc_xHiy`

```{r}
Calc_xHiy <- function(Y, xHi){
#void Calc_xHiy (const gsl_matrix *Y, const gsl_matrix *xHi, gsl_vector *xHiy)
#{
#	gsl_vector_set_zero (xHiy);

#	size_t n_size=Y->size2, d_size=Y->size1, dc_size=xHi->size1;
  n_size <- ncol(Y)
  d_size <- nrow(Y)
  dc_size <- nrow(xHi)
   
  xHiy <- vector(data = 0, length = dc_size)
	#for (size_t k=0; k<n_size; k++) {
  for (k in 1:n_size){
    #	gsl_matrix_const_view xHi_k=gsl_matrix_const_submatrix(xHi, 0, k*d_size, dc_size, d_size);
		xHi_k <- xHi[1:dc_size, ((k - 1) * d_size + 1):(k * d_size)]
    #gsl_vector_const_view y_k=gsl_matrix_const_column(Y, k);
		y_k <- Y[, k]

		#gsl_blas_dgemv (CblasNoTrans, 1.0, &xHi_k.matrix, &y_k.vector, 1.0, xHiy);
		xHiy <- xHiy + xHi_k %*% y_k
		xHi_k -> xHi[1:dc_size, ((k - 1) * d_size + 1):(k * d_size)]
    #gsl_vector_const_view y_k=gsl_matrix_const_column(Y, k);
		y_k -> Y[, k]
		}

#	return;
  return(xHiy)
  }
```


### `Calc_yHiy`

```{r}
#double Calc_yHiy (const gsl_matrix *Y, const gsl_matrix *Hiy_all)
#{
Calc_yHiy <- function(Y, Hiy_all){
	#
  #double yHiy=0.0, d;
	yHiy <- 0
  #size_t n_size=Y->size2;
	n_size <- ncol(Y)

	#for (size_t k=0; k<n_size; k++) {
	for (k in 1:n_size){
	  y_k <- Y[, k]
		#gsl_vector_const_view y_k=gsl_matrix_const_column(Y, k);
		#gsl_vector_const_view Hiy_k=gsl_matrix_const_column(Hiy_all, k);
	  Hiy_k <- Hiy_all[, k]

		#gsl_blas_ddot (&Hiy_k.vector, &y_k.vector, &d);
		d <- Hiy_k %*% y_k
	  #yHiy+=d;
		yHiy <- yHiy + d
		y_k -> Y[, k]
		Hiy_k -> Hiy_all[, k]

	} 
	return(yHiy)

	#return yHiy;
}

```


### `CalcDev`

According to Zhou's comment, this is where the derivatives are calculated.

```{r}
#void CalcDev (const char func_name, const gsl_vector *eval, const gsl_matrix *Qi, const gsl_matrix *Hi, const gsl_matrix *xHi, const gsl_matrix *Hiy, const gsl_vector *QixHiy, gsl_vector *gradient, gsl_matrix *Hessian_inv, double &crt_a, double &crt_b, double &crt_c)
#{
CalcDev <- function(func_name = "R", eval, Qi, Hi, xHi, Hiy, QixHiy){
	#if (func_name!='R' && func_name!='L' && func_name!='r' && func_name!='l') {cout<<"func_name only takes 'R' or 'L': 'R' for log-restricted likelihood, 'L' for log-likelihood."<<endl; return;}

	#size_t dc_size=Qi->size1, d_size=Hi->size1;
	dc_size <- nrow(Qi)
	d_size <- nrow(Hi)
  #size_t c_size=dc_size/d_size;
	c_size <- dc_size / d_size
	
	#size_t v_size=d_size*(d_size+1)/2;
	v_size <- d_size * (d_size + 1) / 2
	#size_t v1, v2;
	#double dev1_g, dev1_e, dev2_gg, dev2_ee, dev2_ge;

	#gsl_matrix *Hessian=gsl_matrix_alloc (v_size*2, v_size*2);
  Hessian <- matrix(nrow = v_size * 2, ncol = v_size * 2)
	#gsl_matrix *xHiDHiy_all_g=gsl_matrix_alloc (dc_size, v_size);
  xHiDHiy_ll_g <- matrix(nrow = dc_size, ncol = v_size)
	#gsl_matrix *xHiDHiy_all_e=gsl_matrix_alloc (dc_size, v_size);
  xHiDHiy_all_e <- matrix(nrow = dc_size, ncol = v_size)
	#gsl_matrix *xHiDHix_all_g=gsl_matrix_alloc (dc_size, v_size*dc_size);
  xHiDHix_all_g <- matrix(nrow = dc_size, ncol = v_size * dc_size)
	#gsl_matrix *xHiDHix_all_e=gsl_matrix_alloc (dc_size, v_size*dc_size);
  xHiDHix_all_e <- matrix(nrow = dc_size, ncol = v_size * dc_size)
	#gsl_matrix *xHiDHixQixHiy_all_g=gsl_matrix_alloc (dc_size, v_size);
  xHiDHixQixHiy_all_g <- matrix(nrow = dc_size, ncol = v_size)
	#gsl_matrix *xHiDHixQixHiy_all_e=gsl_matrix_alloc (dc_size, v_size);
  xHiDHixQixHiy_all_e <- matrix(nrow = dc_size, ncol = v_size)

	#gsl_matrix *QixHiDHiy_all_g=gsl_matrix_alloc (dc_size, v_size);
  QixHiDHiy_all_g <- matrix(nrow = dc_size, ncol = v_size)
	#gsl_matrix *QixHiDHiy_all_e=gsl_matrix_alloc (dc_size, v_size);
  QixHiDHiy_all_e <- matrix(nrow = dc_size, ncol = v_size)
	#gsl_matrix *QixHiDHix_all_g=gsl_matrix_alloc (dc_size, v_size*dc_size);
  QixHiDHix_all_g <- matrix(nrow = dc_size, ncol = v_size * dc_size)
	#gsl_matrix *QixHiDHix_all_e=gsl_matrix_alloc (dc_size, v_size*dc_size);
  QixHiDHix_all_e <- matrix(nrow = dc_size, ncol = v_size * dc_size)
	#gsl_matrix *QixHiDHixQixHiy_all_g=gsl_matrix_alloc (dc_size, v_size);
  QixHiDHixQixHiy_all_g <- matrix(nrow = dc_size, ncol = v_size)
	#gsl_matrix *QixHiDHixQixHiy_all_e=gsl_matrix_alloc (dc_size, v_size);
  QixHiDHixQixHiy_all_e <- matrix(nrow = dc_size, ncol = v_size)

	#gsl_matrix *xHiDHiDHiy_all_gg=gsl_matrix_alloc (dc_size, v_size*v_size);
  xHiDHiDHiy_all_gg <- matrix(nrow = dc_size, ncol = v_size^2)
	#gsl_matrix *xHiDHiDHiy_all_ee=gsl_matrix_alloc (dc_size, v_size*v_size);
  xHiDHiDHiy_all_ee <- matrix(nrow = dc_size, ncol = v_size^2)
	#gsl_matrix *xHiDHiDHiy_all_ge=gsl_matrix_alloc (dc_size, v_size*v_size);
  xHiDHiDHiy_all_ge <- matrix(nrow = dc_size, ncol = v_size^2)
	#gsl_matrix *xHiDHiDHix_all_gg=gsl_matrix_alloc (dc_size, v_size*v_size*dc_size);
  xHiDHiDHix_all_gg <- matrix(nrow = dc_size, ncol = v_size^2 * dc_size)
	#gsl_matrix *xHiDHiDHix_all_ee=gsl_matrix_alloc (dc_size, v_size*v_size*dc_size);
  xHiDHiDHix_all_ee <- matrix(nrow = dc_size, ncol = v_size^2 * dc_size)
	#gsl_matrix *xHiDHiDHix_all_ge=gsl_matrix_alloc (dc_size, v_size*v_size*dc_size);#
  xHiDHiDHix_all_ge <- matrix(nrow = dc_size, ncol = v_size^2 * dc_size)

	#//calculate xHiDHiy_all, xHiDHix_all and xHiDHixQixHiy_all
# 	Calc_xHiDHiy_all (eval, xHi, Hiy, xHiDHiy_all_g, xHiDHiy_all_e);
  foo <- Calc_xHiDHiy_all (eval, xHi, Hiy)
  foo[[1]] -> xHiDHiy_all_g
  foo[[2]] -> xHiDHiy_all_e
# 	Calc_xHiDHix_all (eval, xHi, xHiDHix_all_g, xHiDHix_all_e);
  foo <- Calc_xHiDHix_all (eval, xHi)
  foo[[1]] -> xHiDHix_all_g
  foo[[2]] -> xHiDHix_all_e
# 	Calc_xHiDHixQixHiy_all (xHiDHix_all_g, xHiDHix_all_e, QixHiy, xHiDHixQixHiy_all_g, xHiDHixQixHiy_all_e);
  foo <- Calc_xHiDHixQixHiy_all (xHiDHix_all_g, xHiDHix_all_e, QixHiy)
  xHiDHixQixHiy_all_g <- foo[[1]]
  xHiDHixQixHiy_all_e <- foo[[2]]
# 
# 	Calc_xHiDHiDHiy_all (v_size, eval, Hi, xHi, Hiy, xHiDHiDHiy_all_gg, xHiDHiDHiy_all_ee, xHiDHiDHiy_all_ge);
  foo <- Calc_xHiDHiDHiy_all(v_size, eval, Hi, xHi, Hiy)
  foo[[1]] -> xHiDHiDHiy_all_gg
  foo[[2]] -> xHiDHiDHiy_all_ee
  foo[[3]] -> xHiDHiDHiy_all_ge
# 	Calc_xHiDHiDHix_all (v_size, eval, Hi, xHi, xHiDHiDHix_all_gg, xHiDHiDHix_all_ee, xHiDHiDHix_all_ge);
  foo <- Calc_xHiDHiDHix_all (v_size, eval, Hi, xHi)
  xHiDHiDHix_all_gg <- foo[[1]]
  xHiDHiDHix_all_ee <- foo[[2]]
  xHiDHiDHix_all_ge <- foo[[3]]
# 
# 	//calculate QixHiDHiy_all, QixHiDHix_all and QixHiDHixQixHiy_all
# 	Calc_QiVec_all (Qi, xHiDHiy_all_g, xHiDHiy_all_e, QixHiDHiy_all_g, QixHiDHiy_all_e);
  foo <- Calc_QiVec_all (Qi, xHiDHiy_all_g, xHiDHiy_all_e)
  foo[[1]] -> QixHiDHiy_all_g
  foo[[2]] -> QixHiDHiy_all_e
  
# 	Calc_QiVec_all (Qi, xHiDHixQixHiy_all_g, xHiDHixQixHiy_all_e, QixHiDHixQixHiy_all_g, QixHiDHixQixHiy_all_e);
  foo <- Calc_QiVec_all (Qi, xHiDHixQixHiy_all_g, xHiDHixQixHiy_all_e)
  foo[[1]] -> QixHiDHixQixHiy_all_g
  foo[[2]] -> QixHiDHixQixHiy_all_e
# 	Calc_QiMat_all (Qi, xHiDHix_all_g, xHiDHix_all_e, QixHiDHix_all_g, QixHiDHix_all_e);
  foo <- Calc_QiMat_all (Qi, xHiDHix_all_g, xHiDHix_all_e)
  foo[[1]] -> QixHiDHix_all_g
  foo[[2]] -> QixHiDHix_all_e
# 
# 	double tHiD_g, tHiD_e, tPD_g, tPD_e, tHiDHiD_gg, tHiDHiD_ee, tHiDHiD_ge, tPDPD_gg, tPDPD_ee, tPDPD_ge;
# 	double yPDPy_g, yPDPy_e, yPDPDPy_gg, yPDPDPy_ee, yPDPDPy_ge;
# 
# 	#//calculate gradient and Hessian for Vg
# 	for (size_t i1=0; i1<d_size; i1++) {
  for (i1 in 1:d_size){
# 		for (size_t j1=0; j1<d_size; j1++) {
    for (j1 in 1:d_size){
# 			if (j1<i1) {continue;}
      if (!(j1 < i1)){
        v1 <- GetIndex (i1, j1, d_size)
        Calc_yPDPy(eval, Hiy, QixHiy, xHiDHiy_all_g, xHiDHiy_all_e, xHiDHixQixHiy_all_g, xHiDHixQixHiy_all_e, i1, j1)
      
# 			v1=GetIndex (i1, j1, d_size);
# 
# 			Calc_yPDPy (eval, Hiy, QixHiy, xHiDHiy_all_g, xHiDHiy_all_e, xHiDHixQixHiy_all_g, xHiDHixQixHiy_all_e, i1, j1, yPDPy_g, yPDPy_e);
# 
# 			if (func_name=='R' || func_name=='r') {
# 				Calc_tracePD (eval, Qi, Hi, xHiDHix_all_g, xHiDHix_all_e, i1, j1, tPD_g, tPD_e);
        foo <- Calc_tracePD(eval, Qi, Hi, xHiDHix_all_g, xHiDHix_all_e, i1, j1)
        foo[[1]] -> tPD_g
        foo[[2]] -> tPD_e
        #//cout<<i1<<" "<<j1<<" "<<yPDPy_g<<" "<<yPDPy_e<<" "<<tPD_g<<" "<<tPD_e<<endl;
# 
# 				dev1_g=-0.5*tPD_g+0.5*yPDPy_g;
        dev1_g <- - 0.5 * tPD_g + 0.5 * yPDPy_g
# 				dev1_e=-0.5*tPD_e+0.5*yPDPy_e;
        dev1_e <- - 0.5 * tPD_e + 0.5 * yPDPy_e
# 			} else {
# 				Calc_traceHiD (eval, Hi, i1, j1, tHiD_g, tHiD_e);
# 
# 				dev1_g=-0.5*tHiD_g+0.5*yPDPy_g;
# 				dev1_e=-0.5*tHiD_e+0.5*yPDPy_e;
# 			}
# 
# 			gsl_vector_set (gradient, v1, dev1_g);
        gradient[v1] <- dev1_g
# 			gsl_vector_set (gradient, v1+v_size, dev1_e);
        gradient[v1 + v_size] <- dev1_e
# 
# 			for (size_t i2=0; i2<d_size; i2++) {
        for (i2 in 1:d_size){
# 				for (size_t j2=0; j2<d_size; j2++) {
          for (j2 in 1:d_size){
# 					if (j2<i2) {continue;}
            if (!(j2 < i2)){
# 					v2=GetIndex (i2, j2, d_size);
              v2 <- GetIndex(i2, j2, d_size)
# 
# 					if (v2<v1) {continue;}
              if (!(v2 < v1)){
                
              
# 
# 					Calc_yPDPDPy (eval, Hi, xHi, Hiy, QixHiy, xHiDHiy_all_g, xHiDHiy_all_e, QixHiDHiy_all_g, QixHiDHiy_all_e, xHiDHixQixHiy_all_g, xHiDHixQixHiy_all_e, QixHiDHixQixHiy_all_g, QixHiDHixQixHiy_all_e, xHiDHiDHiy_all_gg, xHiDHiDHiy_all_ee, xHiDHiDHiy_all_ge, xHiDHiDHix_all_gg, xHiDHiDHix_all_ee, xHiDHiDHix_all_ge, i1, j1, i2, j2, yPDPDPy_gg, yPDPDPy_ee, yPDPDPy_ge);
                foo <- Calc_yPDPDPy (eval, Hi, xHi, Hiy, QixHiy, xHiDHiy_all_g, xHiDHiy_all_e, QixHiDHiy_all_g, QixHiDHiy_all_e, xHiDHixQixHiy_all_g, xHiDHixQixHiy_all_e, QixHiDHixQixHiy_all_g, QixHiDHixQixHiy_all_e, xHiDHiDHiy_all_gg, xHiDHiDHiy_all_ee, xHiDHiDHiy_all_ge, xHiDHiDHix_all_gg, xHiDHiDHix_all_ee, xHiDHiDHix_all_ge, i1, j1, i2, j2)
                foo[[1]] -> yPDPDPy_gg
                foo[[2]] -> yPDPDPy_ee
                foo[[3]] -> yPDPDPy_ge
# 
# 					//cout<<i1<<" "<<j1<<" "<<i2<<" "<<j2<<" "<<yPDPDPy_gg<<" "<<yPDPDPy_ee<<" "<<yPDPDPy_ge<<endl;
# 					//AI for reml
# 					if (func_name=='R' || func_name=='r') {
# 						Calc_tracePDPD (eval, Qi, Hi, xHi, QixHiDHix_all_g, QixHiDHix_all_e, xHiDHiDHix_all_gg, xHiDHiDHix_all_ee, xHiDHiDHix_all_ge, i1, j1, i2, j2, tPDPD_gg, tPDPD_ee, tPDPD_ge);
                foo <- Calc_tracePDPD (eval, Qi, Hi, xHi, QixHiDHix_all_g, QixHiDHix_all_e, xHiDHiDHix_all_gg, xHiDHiDHix_all_ee, xHiDHiDHix_all_ge, i1, j1, i2, j2)
                foo[[1]] -> tPDPD_gg
                foo[[2]] -> tPDPD_ee
                foo[[3]] -> tPDPD_ge
# 
# 						dev2_gg=0.5*tPDPD_gg-yPDPDPy_gg;
                dev2_gg <- 0.5 * tPDPD_gg-yPDPDPy_gg
# 						dev2_ee=0.5*tPDPD_ee-yPDPDPy_ee;
                dev2_ee <- 0.5 * tPDPD_ee - yPDPDPy_ee
# 						dev2_ge=0.5*tPDPD_ge-yPDPDPy_ge;
                dev2_ge <- 0.5 * tPDPD_ge - yPDPDPy_ge
# 						/*
# 						dev2_gg=-0.5*yPDPDPy_gg;
# 						dev2_ee=-0.5*yPDPDPy_ee;
# 						dev2_ge=-0.5*yPDPDPy_ge;
# 						*/
# 					} else {
# 						Calc_traceHiDHiD (eval, Hi, i1, j1, i2, j2, tHiDHiD_gg, tHiDHiD_ee, tHiDHiD_ge);
# 
# 						dev2_gg=0.5*tHiDHiD_gg-yPDPDPy_gg;
# 						dev2_ee=0.5*tHiDHiD_ee-yPDPDPy_ee;
# 						dev2_ge=0.5*tHiDHiD_ge-yPDPDPy_ge;
# 					}
# 
# 					//set up Hessian
# 					gsl_matrix_set (Hessian, v1, v2, dev2_gg);
                Hessian[v1, v2] <- dev2_gg
# 					gsl_matrix_set (Hessian, v1+v_size, v2+v_size, dev2_ee);
                Hessian[v1 + v_size, v2 + v_size] <- dev2_ee
# 					gsl_matrix_set (Hessian, v1, v2+v_size, dev2_ge);
                Hessian[v1, v2 + v_size] <- dev2_ge
# 					gsl_matrix_set (Hessian, v2+v_size, v1, dev2_ge);
                Hessian[v2 + v_size, v1] <- dev2_ge
# 
# 					if (v1!=v2) {
                if (v1 != v2){
# 						gsl_matrix_set (Hessian, v2, v1, dev2_gg);
                  Hessian[v2, v1] <- dev2_gg
# 						gsl_matrix_set (Hessian, v2+v_size, v1+v_size, dev2_ee);
                  Hessian[v2 + v_size, v1 + v_size] <- dev2_ee
# 						gsl_matrix_set (Hessian, v2, v1+v_size, dev2_ge);
                  Hessian[v2, v1 + v_size] <- dev2_ge
# 						gsl_matrix_set (Hessian, v1+v_size, v2, dev2_ge);
                  Hessian[v1 + v_size, v2] <- dev2_ge
                }
# 					}
              }
            }
# 				}
      } # end if 
    } # end j loop
  } # end i loop
    
# 			}
# 		}
# 	}
# 
# 	/*
# 	cout<<"Hessian: "<<endl;
# 	for (size_t i=0; i<2*v_size; i++) {
# 		for (size_t j=0; j<2*v_size; j++) {
# 			cout<<gsl_matrix_get(Hessian, i, j)<<"\t";
# 		}
# 		cout<<endl;
# 	}
# 	*/
# 
# 
# 	//Invert Hessian
# 	int sig;
# 	gsl_permutation * pmt=gsl_permutation_alloc (v_size*2);
# 
# 	LUDecomp (Hessian, pmt, &sig);
# 	LUInvert (Hessian, pmt, Hessian_inv);
  Hessian_inv <- solve(Hessian)

  # 	/*
# 	cout<<"Hessian Inverse: "<<endl;
# 	for (size_t i=0; i<2*v_size; i++) {
# 		for (size_t j=0; j<2*v_size; j++) {
# 			cout<<gsl_matrix_get(Hessian_inv, i, j)<<"\t";
# 		}
# 		cout<<endl;
# 	}
# 	*/
# 	gsl_permutation_free(pmt);
# 	gsl_matrix_free(Hessian);
# 
# 	//calculate Edgeworth correction factors
# 	//after inverting Hessian
# 	if (c_size>1) {
# 		CalcCRT (Hessian_inv, Qi, QixHiDHix_all_g, QixHiDHix_all_e, xHiDHiDHix_all_gg, xHiDHiDHix_all_ee, xHiDHiDHix_all_ge, d_size, crt_a, crt_b, crt_c);
# 	} else {
# 		crt_a=0.0; crt_b=0.0; crt_c=0.0;
# 	}
# 
# 	gsl_matrix_free(xHiDHiy_all_g);
# 	gsl_matrix_free(xHiDHiy_all_e);
# 	gsl_matrix_free(xHiDHix_all_g);
# 	gsl_matrix_free(xHiDHix_all_e);
# 	gsl_matrix_free(xHiDHixQixHiy_all_g);
# 	gsl_matrix_free(xHiDHixQixHiy_all_e);
# 
# 	gsl_matrix_free(QixHiDHiy_all_g);
# 	gsl_matrix_free(QixHiDHiy_all_e);
# 	gsl_matrix_free(QixHiDHix_all_g);
# 	gsl_matrix_free(QixHiDHix_all_e);
# 	gsl_matrix_free(QixHiDHixQixHiy_all_g);
# 	gsl_matrix_free(QixHiDHixQixHiy_all_e);
# 
# 	gsl_matrix_free(xHiDHiDHiy_all_gg);
# 	gsl_matrix_free(xHiDHiDHiy_all_ee);
# 	gsl_matrix_free(xHiDHiDHiy_all_ge);
# 	gsl_matrix_free(xHiDHiDHix_all_gg);
# 	gsl_matrix_free(xHiDHiDHix_all_ee);
# 	gsl_matrix_free(xHiDHiDHix_all_ge);
# 
# 	return;
# }
    }
  }
  return(list(Hessian_inv, gradient))
}
```


### `Calc_xHiDHiy_all`

```{r}
#void Calc_xHiDHiy_all (const gsl_vector *eval, const gsl_matrix *xHi, const gsl_matrix *Hiy, gsl_matrix *xHiDHiy_all_g, gsl_matrix *xHiDHiy_all_e)
#{
Calc_xHiDHiy_all <- function(eval, xHi, Hiy){
 	#gsl_matrix_set_zero(xHiDHiy_all_g);
	#gsl_matrix_set_zero(xHiDHiy_all_e);
  
	#size_t d_size=Hiy->size1;
	#size_t v;
  d_size <- nrow(Hiy)
	#for (size_t i=0; i<d_size; i++) {
	for (i in 1:d_size){
  #	for (size_t j=0; j<d_size; j++) {
	  for (j in 1:d_size){
			#if (j<i) {continue;}
			if (j >= i){
	    #v=GetIndex(i, j, d_size);
			  v <- GetIndex(i, j, d_size)
      
			#gsl_vector_view xHiDHiy_g=gsl_matrix_column (xHiDHiy_all_g, v);
			  xHiDHiy_g <- xHiDHiy_all_g[, v]
			#gsl_vector_view xHiDHiy_e=gsl_matrix_column (xHiDHiy_all_e, v);
			  xHiDHiy_e <- xHiDHiy_all_e[, v]

			#Calc_xHiDHiy (eval, xHi, Hiy, i, j, &xHiDHiy_g.vector, &xHiDHiy_e.vector);
			  Calc_xHiDHiy(eval, xHi, Hiy, i, j, xHiDHiy_g, xHiDHiy_e)
			  xHiDHiy_g -> xHiDHiy_all_g[, v]
			  xHiDHiy_e -> xHiDHiy_all_e[, v]
			}
		}
	}
	#return;
	  return(list(xHiDHiy_all_g, xHiDHiy_all_e))
}

```



### `Calc_xHiDHix_all`

```{r}
#//calculate (xHiDHix) for every pair of i j
#void Calc_xHiDHix_all (const gsl_vector *eval, const gsl_matrix *xHi, gsl_matrix *xHiDHix_all_g, gsl_matrix *xHiDHix_all_e)
#{
Calc_xHiDHix_all <- function(eval, xHi){
  
	#gsl_matrix_set_zero(xHiDHix_all_g);
	#gsl_matrix_set_zero(xHiDHix_all_e);
  

	#size_t d_size=xHi->size2/eval->size, dc_size=xHi->size1;
	#size_t v;
  d_size <- ncol(xHi) / length(eval)
  dc_size <- nrow(xHi)

	#for (size_t i=0; i<d_size; i++) {
  for (i in 1:d_size){
	#	for (size_t j=0; j<d_size; j++) {
    for (j in 1:d_size){
			#if (j<i) {continue;}
      if (i < j){
        
			#v=GetIndex(i, j, d_size);
        v <- GetIndex(i, j, d_size)
			#gsl_matrix_view xHiDHix_g=gsl_matrix_submatrix (xHiDHix_all_g, 0, v*dc_size, dc_size, dc_size);
			#gsl_matrix_view xHiDHix_e=gsl_matrix_submatrix (xHiDHix_all_e, 0, v*dc_size, dc_size, dc_size);
        xHiDHix_g <- xHiDHix_all_g[1:dc_size, (v * (dc_size - 1) + 1):(v * dc_size)]
        xHiDHix_e <- xHiDHix_all_e[1:dc_size, (v * (dc_size - 1) + 1):(v * dc_size)]
			#Calc_xHiDHix (eval, xHi, i, j, &xHiDHix_g.matrix, &xHiDHix_e.matrix);
        Calc_xHiDHix(eval, xHi, i, j, xHiDHix_g, xHiDHix_e)
        xHiDHix_g -> xHiDHix_all_g[1:dc_size, (v * (dc_size - 1) + 1):(v * dc_size)]
        xHiDHix_e -> xHiDHix_all_e[1:dc_size, (v * (dc_size - 1) + 1):(v * dc_size)]
      }
		}
	}
	#return;
  return(list(xHiDHix_all_g, xHiDHix_all_e))
}
```





### `Calc_xHiDHixQixHiy_all`

```{r}
#void Calc_xHiDHixQixHiy_all (const gsl_matrix *xHiDHix_all_g, const gsl_matrix *xHiDHix_all_e, const gsl_vector *QixHiy, gsl_matrix *xHiDHixQixHiy_all_g, gsl_matrix *xHiDHixQixHiy_all_e)
#{
Calc_xHiDHixQixHiy_all <- function(xHiDHix_all_g, xHiDHix_all_e, QixHiy, xHiDHixQixHiy_all_g, xHiDHixQixHiy_all_e){
	dc_size <- nrow(xHiDHix_all_g)
  #size_t dc_size=xHiDHix_all_g->size1;
	#size_t v_size=xHiDHix_all_g->size2/dc_size;
	v_size <- ncol(xHiDHix_all_g) / dc_size

	#for (size_t i=0; i<v_size; i++) {
	for (i in 1:v_size){
		#gsl_matrix_const_view xHiDHix_g=gsl_matrix_const_submatrix (xHiDHix_all_g, 0, i*dc_size, dc_size, dc_size);
		#gsl_matrix_const_view xHiDHix_e=gsl_matrix_const_submatrix (xHiDHix_all_e, 0, i*dc_size, dc_size, dc_size);
	  xHiDHix_g <- xHiDHix_all_g[1:dc_size, ((i - 1) * dc_size + 1):(i * dc_size)]
	  xHiDHix_e <- xHiDHix_all_e[1:dc_size, ((i - 1) * dc_size + 1):(i * dc_size)]

		#gsl_vector_view xHiDHixQixHiy_g=gsl_matrix_column (xHiDHixQixHiy_all_g, i);
		#gsl_vector_view xHiDHixQixHiy_e=gsl_matrix_column (xHiDHixQixHiy_all_e, i);
	  xHiDHixQixHiy_g <- xHiDHixQixHiy_all_g[, i]
	  xHiDHixQixHiy_e <- xHiDHixQixHiy_all_e[, i]

		#gsl_blas_dgemv (CblasNoTrans, 1.0, &xHiDHix_g.matrix, QixHiy, 0.0, &xHiDHixQixHiy_g.vector);
		xHiDHix_g %*% QixHiy -> xHiDHixQixHiy_g
	  #gsl_blas_dgemv (CblasNoTrans, 1.0, &xHiDHix_e.matrix, QixHiy, 0.0, &xHiDHixQixHiy_e.vector);
		xHiDHix_e %*% QixHiy -> xHiDHixQixHiy_e
	  xHiDHixQixHiy_g -> xHiDHixQixHiy_all_g[, i]
	  xHiDHixQixHiy_e -> xHiDHixQixHiy_all_e[, i]

	}
	return(list(xHiDHixQixHiy_all_g, xHiDHixQixHiy_all_e))
	#return;
}
```




### `Calc_xHiDHiDHiy_all`

```{r}
Calc_xHiDHiDHiy_all <- function(v_size, eval, Hi, xHi, Hiy){
#void Calc_xHiDHiDHiy_all (const size_t v_size, const gsl_vector *eval, const gsl_matrix *Hi, const gsl_matrix *xHi, const gsl_matrix *Hiy, gsl_matrix *xHiDHiDHiy_all_gg, gsl_matrix *xHiDHiDHiy_all_ee, gsl_matrix *xHiDHiDHiy_all_ge)
#{
	#gsl_matrix_set_zero(xHiDHiDHiy_all_gg);
	#gsl_matrix_set_zero(xHiDHiDHiy_all_ee);
	#gsl_matrix_set_zero(xHiDHiDHiy_all_ge);

	#size_t d_size=Hiy->size1;
	d_size <- nrow(Hiy)
  #size_t v1, v2;

	
	#for (size_t i1=0; i1<d_size; i1++) {
	for (i1 in 1:d_size){
	  	#for (size_t j1=0; j1<d_size; j1++) {
	  for (j1 in 1:d_size){
	    if (i1 < j1){
			#if (j1<i1) {continue;}
	      v1 <- GetIndex(i1, j1, d_size)
			
	      #v1=GetIndex(i1, j1, d_size);

			#for (size_t i2=0; i2<d_size; i2++) {
	      for (i2 in 1:d_size){
	        for (j2 in 1:d_size){
				#for (size_t j2=0; j2<d_size; j2++) {
					#if (j2<i2) {continue;}
	          if (i2 < j2){
					#v2=GetIndex(i2, j2, d_size);
	            v2 <- GetIndex(i2, j2, d_size)

					#gsl_vector_view xHiDHiDHiy_gg=gsl_matrix_column (xHiDHiDHiy_all_gg, v1*v_size+v2);
	            xHiDHiDHiy_gg <- xHiDHiDHiy_all_gg[, (v1 - 1) * v_size + v2] # CHECK INDICES HERE
	            xHiDHiDHiy_ee <- xHiDHiDHiy_all_ee[, (v1 - 1) * v_size + v2] # CHECK INDICES HERE
	            xHiDHiDHiy_ge <- xHiDHiDHiy_all_ge[, (v1 - 1) * v_size + v2] # CHECK INDICES HERE
	            #gsl_vector_view xHiDHiDHiy_ee=gsl_matrix_column (xHiDHiDHiy_all_ee, v1*v_size+v2);
					#gsl_vector_view xHiDHiDHiy_ge=gsl_matrix_column (xHiDHiDHiy_all_ge, v1*v_size+v2);

	            Calc_xHiDHiDHiy (eval, Hi, xHi, Hiy, i1, j1, i2, j2, xHiDHiDHiy_gg, xHiDHiDHiy_ee, xHiDHiDHiy_ge)
					
	            #Calc_xHiDHiDHiy (eval, Hi, xHi, Hiy, i1, j1, i2, j2, &xHiDHiDHiy_gg.vector, &xHiDHiDHiy_ee.vector, &xHiDHiDHiy_ge.vector);
	            xHiDHiDHiy_gg -> xHiDHiDHiy_all_gg[, (v1 - 1) * v_size + v2] # CHECK INDICES HERE
	            xHiDHiDHiy_ee -> xHiDHiDHiy_all_ee[, (v1 - 1) * v_size + v2] # CHECK INDICES HERE
	            xHiDHiDHiy_ge -> xHiDHiDHiy_all_ge[, (v1 - 1) * v_size + v2] # CHECK INDICES HERE
	          }
	        }
				}
			}
		}
	}
	#return;
	return(list(xHiDHiDHiy_all_gg, xHiDHiDHiy_all_ee, xHiDHiDHiy_all_ge))
}
```




### `Calc_xHiDHiDHix_all`

```{r}
Calc_xHiDHiDHix_all <- function(v_size, eval, Hi, xHi){
#   //calculate (xHiDHix) for every pair of i j
# void Calc_xHiDHiDHix_all (const size_t v_size, const gsl_vector *eval, const gsl_matrix *Hi, const gsl_matrix *xHi, gsl_matrix *xHiDHiDHix_all_gg, gsl_matrix *xHiDHiDHix_all_ee, gsl_matrix *xHiDHiDHix_all_ge)
# {
# 	gsl_matrix_set_zero(xHiDHiDHix_all_gg);
# 	gsl_matrix_set_zero(xHiDHiDHix_all_ee);
# 	gsl_matrix_set_zero(xHiDHiDHix_all_ge);
# 
  
# 	size_t d_size=xHi->size2/eval->size, dc_size=xHi->size1;
# 	size_t v1, v2;
# 
  d_size <- ncol(xHi) / length(eva)
  dc_size <- nrow(xHi)
  
# 	for (size_t i1=0; i1<d_size; i1++) {
  for (i1 in 1:d_size){
# 		for (size_t j1=0; j1<d_size; j1++) {
    for (j1 in 1:d_size){

    # 			if (j1<i1) {continue;}
      if (i1 <= j1){
# 			v1=GetIndex(i1, j1, d_size);
# 
# 			for (size_t i2=0; i2<d_size; i2++) {
        for (i2 in 1:d_size){
          
        
# 				for (size_t j2=0; j2<d_size; j2++) {
          for (j2 in 1:d_size){
# 			
          #if (j2<i2) {continue;}
 				
            if (i2 <= j2){
            #v2=GetIndex(i2, j2, d_size);
              v2 <- GetIndex(i2, j2, d_size)
# 
# 					if (v2<v1) {continue;}
              if (!(v2 < v1)){
# 
# 					gsl_matrix_view xHiDHiDHix_gg1=gsl_matrix_submatrix (xHiDHiDHix_all_gg, 0, (v1*v_size+v2)*dc_size, dc_size, dc_size);
                xHiDHiDHix_gg1 <- xHiDHiDHix_all_gg[1:dc_size, (((v1 - 1) * v_size + v2 - 1) * dc_size + 1):(((v1 - 1) * v_size + v2 - 1) * dc_size)]
                                
# 					gsl_matrix_view xHiDHiDHix_ee1=gsl_matrix_submatrix (xHiDHiDHix_all_ee, 0, (v1*v_size+v2)*dc_size, dc_size, dc_size);
                xHiDHiDHix_ee1 <- xHiDHiDHix_all_ee[1:dc_size, (((v1 - 1) * v_size + v2 - 1) * dc_size + 1):(((v1 - 1) * v_size + v2 - 1) * dc_size)]
                                
# 					gsl_matrix_view xHiDHiDHix_ge1=gsl_matrix_submatrix (xHiDHiDHix_all_ge, 0, (v1*v_size+v2)*dc_size, dc_size, dc_size);
                xHiDHiDHix_ge1 <- xHiDHiDHix_all_ge[1:dc_size, (((v1 - 1) * v_size + v2 - 1) * dc_size + 1):(((v1 - 1) * v_size + v2 - 1) * dc_size)]
                                
# 
# 					Calc_xHiDHiDHix (eval, Hi, xHi, i1, j1, i2, j2, &xHiDHiDHix_gg1.matrix, &xHiDHiDHix_ee1.matrix, &xHiDHiDHix_ge1.matrix);
                cx_out <- Calc_xHiDHiDHix() # ADD ARGUMENTS!
# 
# 					if (v2!=v1) {
                if (v2 != v1){
                  xHiDHiDHix_gg2 <- xHiDHiDHix_all_gg[1:dc_size, (((v2 - 1)* v_size + v1 - 1) * dc_size + 1):(((v2 - 1)* v_size + v1) * dc_size)]
                  xHiDHiDHix_ee2 <- xHiDHiDHix_all_ee[1:dc_size, (((v2 - 1)* v_size + v1 - 1) * dc_size + 1):(((v2 - 1)* v_size + v1) * dc_size)]
                  xHiDHiDHix_ge2 <- xHiDHiDHix_all_ge[1:dc_size, (((v2 - 1)* v_size + v1 - 1) * dc_size + 1):(((v2 - 1)* v_size + v1) * dc_size)]
                  
                  xHiDHiDHix_gg2 <- xHiDHiDHix_gg1
                  xHiDHiDHix_ee2 <- xHiDHiDHix_ee1
                  xHiDHiDHix_ge2 <- xHiDHiDHix_ge1
                  
                  xHiDHiDHix_gg2 -> xHiDHiDHix_all_gg[1:dc_size, (((v2 - 1)* v_size + v1 - 1) * dc_size + 1):(((v2 - 1)* v_size + v1) * dc_size)]
                  xHiDHiDHix_ee2 -> xHiDHiDHix_all_ee[1:dc_size, (((v2 - 1)* v_size + v1 - 1) * dc_size + 1):(((v2 - 1)* v_size + v1) * dc_size)]
                  xHiDHiDHix_ge2 -> xHiDHiDHix_all_ge[1:dc_size, (((v2 - 1)* v_size + v1 - 1) * dc_size + 1):(((v2 - 1)* v_size + v1) * dc_size)]
                  
                }
# 						gsl_matrix_view xHiDHiDHix_gg2=gsl_matrix_submatrix (xHiDHiDHix_all_gg, 0, (v2*v_size+v1)*dc_size, dc_size, dc_size);
# 						gsl_matrix_view xHiDHiDHix_ee2=gsl_matrix_submatrix (xHiDHiDHix_all_ee, 0, (v2*v_size+v1)*dc_size, dc_size, dc_size);
# 						gsl_matrix_view xHiDHiDHix_ge2=gsl_matrix_submatrix (xHiDHiDHix_all_ge, 0, (v2*v_size+v1)*dc_size, dc_size, dc_size);
# 
# 						gsl_matrix_memcpy (&xHiDHiDHix_gg2.matrix, &xHiDHiDHix_gg1.matrix);
# 						gsl_matrix_memcpy (&xHiDHiDHix_ee2.matrix, &xHiDHiDHix_ee1.matrix);
# 						gsl_matrix_memcpy (&xHiDHiDHix_ge2.matrix, &xHiDHiDHix_ge1.matrix);
# 					}
                xHiDHiDHix_gg1 -> xHiDHiDHix_all_gg[1:dc_size, (((v1 - 1) * v_size + v2 - 1) * dc_size + 1):(((v1 - 1) * v_size + v2 - 1) * dc_size)]
                xHiDHiDHix_ee1 -> xHiDHiDHix_all_ee[1:dc_size, (((v1 - 1) * v_size + v2 - 1) * dc_size + 1):(((v1 - 1) * v_size + v2 - 1) * dc_size)]
                xHiDHiDHix_ge1 -> xHiDHiDHix_all_ge[1:dc_size, (((v1 - 1) * v_size + v2 - 1) * dc_size + 1):(((v1 - 1) * v_size + v2 - 1) * dc_size)]
              }
            }
          }
        }
# 				}
# 			}
# 		}
# 	}
#	return;
      } # end if (i1 < j1)
    } # end for loop over j1
  } # end for loop over i1
  return(list(xHiDHiDHix_all_gg, xHiDHiDHix_all_ee, xHiDHiDHix_all_ge))
}
```






### `Calc_QiVec_all`

```{r}
Calc_QiVec_all <- function(Qi, vec_all_g, vec_all_e){
#   //calculate Qi(xHiDHiy) and Qi(xHiDHix)Qi(xHiy) for each pair of i j (i<=j)
# void Calc_QiVec_all (const gsl_matrix *Qi, const gsl_matrix *vec_all_g, const gsl_matrix *vec_all_e, gsl_matrix *Qivec_all_g, gsl_matrix *Qivec_all_e)
# {
# 	for (size_t i=0; i<vec_all_g->size2; i++) {
  for (i in 1:ncol(vec_all_g)){
# 		gsl_vector_const_view vec_g=gsl_matrix_const_column (vec_all_g, i);
    vec_g <- vec_all_g[, i]
    vec_e <- vec_all_e[, i]

    # 		gsl_vector_const_view vec_e=gsl_matrix_const_column (vec_all_e, i);
# 
# 		gsl_vector_view Qivec_g=gsl_matrix_column (Qivec_all_g, i);
    Qivec_g <- Qivec_all_g[, i]
# 		gsl_vector_view Qivec_e=gsl_matrix_column (Qivec_all_e, i);
    Qivec_e <- Qivec_all_e[, i]

    # 
# 		gsl_blas_dgemv (CblasNoTrans, 1.0, Qi, &vec_g.vector, 0.0, &Qivec_g.vector);
    Qivec_g <- Qi %*% vec_g
    Qivec_e <- Qi %*% vec_e
# 		gsl_blas_dgemv (CblasNoTrans, 1.0, Qi, &vec_e.vector, 0.0, &Qivec_e.vector);
# 	}
# 
# 	return;
# }
    vec_g -> vec_all_g[, i]
    vec_e -> vec_all_e[, i]
    Qivec_g -> Qivec_all_g[, i]
    Qivec_e -> Qivec_all_e[, i]
  }
  return(list(Qivec_all_g, Qivec_all_e))
}
```


### `Calc_QiMat_all`

```{r}
# //calculate Qi(xHiDHix) for each pair of i j (i<=j)
# void Calc_QiMat_all (const gsl_matrix *Qi, const gsl_matrix *mat_all_g, const gsl_matrix *mat_all_e, gsl_matrix *Qimat_all_g, gsl_matrix *Qimat_all_e)
# {
Calc_QiMat_all <- function(Qi, mat_all_g, mat_all_e){
  dc_size <- nrow(Qi)
  v_size <- ncol(mat_all_g) / nrow(mat_all_g)
# 	size_t dc_size=Qi->size1;
# 	size_t v_size=mat_all_g->size2/mat_all_g->size1;
# 
# 	for (size_t i=0; i<v_size; i++) {
  for (i in 1:v_size){
    mat_g <- mat_all_g[1:dc_size, ((i - 1) * dc_size + 1):(i * dc_size)]
  # 		gsl_matrix_const_view mat_g=gsl_matrix_const_submatrix (mat_all_g, 0, i*dc_size, dc_size, dc_size);
# 		gsl_matrix_const_view mat_e=gsl_matrix_const_submatrix (mat_all_e, 0, i*dc_size, dc_size, dc_size);
    mat_e <- mat_all_e[1:dc_size, ((i - 1) * dc_size + 1):(i * dc_size)]

    # 
# 		gsl_matrix_view Qimat_g=gsl_matrix_submatrix (Qimat_all_g, 0, i*dc_size, dc_size, dc_size);
    Qimat_g <- Qimat_all_g[1:dc_size, ((i - 1) * dc_size + 1):(i * dc_size)]
# 		gsl_matrix_view Qimat_e=gsl_matrix_submatrix (Qimat_all_e, 0, i*dc_size, dc_size, dc_size);
    Qimat_e <- Qimat_all_e[1:dc_size, ((i - 1) * dc_size + 1):(i * dc_size)]

    # 
# 		gsl_blas_dgemm (CblasNoTrans, CblasNoTrans, 1.0, Qi, &mat_g.matrix, 0.0, &Qimat_g.matrix);
    Qimat_g <- Qi %*% mat_g 
# 		gsl_blas_dgemm (CblasNoTrans, CblasNoTrans, 1.0, Qi, &mat_e.matrix, 0.0, &Qimat_e.matrix);
    Qimat_e <- Qi %*% mat_e
# 	}
    
    mat_g -> mat_all_g[1:dc_size, ((i - 1) * dc_size + 1):(i * dc_size)]
    mat_e -> mat_all_e[1:dc_size, ((i - 1) * dc_size + 1):(i * dc_size)]
    Qimat_g -> Qimat_all_g[1:dc_size, ((i - 1) * dc_size + 1):(i * dc_size)]
    Qimat_e -> Qimat_all_e[1:dc_size, ((i - 1) * dc_size + 1):(i * dc_size)]
    # 
# 	return;
# }
  }
  return(list(Qimat_all_g, Qimat_all_e))
}
```







### `GetIndex`

```{r}
# //0<=i,j<d_size
# size_t GetIndex (const size_t i, const size_t j, const size_t d_size)
# {
GetIndex <- function(i, j, d_size){
# 	if (i>=d_size || j>=d_size) {cout<<"error in GetIndex."<<endl; return 0;}
  if (i >= d_size | j >= d_size) return("error")

  # 
# 	size_t s, l;
# 	if (j<i) {s=j; l=i;} else {s=i; l=j;}
# 
  if (j < i){
    s <- j
    l <- i
  } else {
    s <- i
    l <- j
  }
# 	return (2*d_size-s+1)*s/2+l-s;
  return((2 * d_size - s + 1) * s / 2 + l - s)
# }
}
```


### `Calc_xHiDHiy`

```{r}
Calc_xHiDHiy <- function(eval, xHi, Hiy, i, j){
#   void Calc_xHiDHiy (const gsl_vector *eval, const gsl_matrix *xHi, const gsl_matrix *Hiy, const size_t i, const size_t j, gsl_vector *xHiDHiy_g, gsl_vector *xHiDHiy_e)
# {
# 	gsl_vector_set_zero(xHiDHiy_g);
# 	gsl_vector_set_zero(xHiDHiy_e);
  
# 	size_t n_size=eval->size, d_size=Hiy->size1;
  n_size <- length(eval)
  d_size <- nrow(Hiy)
  xHiDHiy_g <- numeric(length = d_size)
  xHiDHiy_e <- numeric(length = d_size)
# 
# 	double delta, d;
# 
# 	for (size_t k=0; k<n_size; k++) {
  for (k in 1:n_size){
# 		delta=gsl_vector_get (eval, k);
    delta <- eval[k]
# 
# 		gsl_vector_const_view xHi_col_i=gsl_matrix_const_column (xHi, k*d_size+i);
    xHi_col_i <- xHi[, k * d_size + i] #VERIFY INDEX HERE!
# 		d=gsl_matrix_get (Hiy, j, k);
# 
    d <- Hiy[j, k] # VERIFY INDEX HERE!
# 		gsl_blas_daxpy (d*delta, &xHi_col_i.vector, xHiDHiy_g);
    xHiDHiy_g <- xHiDHiy_g + d * delta * xHi_col_i
# 		gsl_blas_daxpy (d, &xHi_col_i.vector, xHiDHiy_e);
    xHiDHiy_e <- xHiDHiy_e + d * xHi_col_i
# 
# 		if (i!=j) {
    if ( i != j){
      
    
# 			gsl_vector_const_view xHi_col_j=gsl_matrix_const_column (xHi, k*d_size+j);
      xHi_col_j <- xHi[ , (k - 1) * d_size + j]
# 			d=gsl_matrix_get (Hiy, i, k);
      d <- Hiy[i, k]
# 
# 			gsl_blas_daxpy (d*delta, &xHi_col_j.vector, xHiDHiy_g);
      xHiDHiy_g <- xhiDHiy_g + d * delta * xHi_col_j
# 			gsl_blas_daxpy (d, &xHi_col_j.vector, xHiDHiy_e);
      xHiDHiy_e <- xhiDHiy_e + d * xHi_col_j
# 		
# 		}
# 	}
# 
# 	return;
# }
    }
  }
  return(list(xHiDHiy_g, xHiDHiy_e))
}
```





