---
title: Using `dplyr` functions in my R package
author: Frederick Boehm
date: '`r lubridate::now()`'
categories:
  - computing
  - packages
  - R
tags:
  - dplyr
slug: using-dplyr-functions-in-my-r-package
draft: 'false'
---

## Background

I wrote the following function for use in my `pleiotropy` R package.

```{r, eval = FALSE}
calc_lod <- function(data, n_mouse){
  # define log10detrss0
  log10detrss0 <- dplyr::filter(data, marker1 == marker2)
  pre1 <- dplyr::group_by(data, marker1)
  profile1 <- dplyr::summarise(pre1, profile = min(log10detrss))
  pre2 <- dplyr::group_by(data, marker2)
  profile2 <- dplyr::summarise(pre2, profile = min(log10detrss))
  return(tibble::tibble(lod1 = - n_mouse * (profile1$profile - min(log10detrss0$log10detrss)) / 2,
                        lod2 = - n_mouse * (profile2$profile - min(log10detrss0$log10detrss))/ 2,
                        joint = - n_mouse * (log10detrss0$log10detrss - min(log10detrss0$log10detrss)) / 2))
}

```

I then wanted to include this function in my R package `pleiotropy`. However, when I run `devtools::check()` for this package, I get a note. Specifically, the note says:

> checking R code for possible problems ... NOTE

> calc_lod: no visible binding for global variable ‘marker1’

> calc_lod: no visible binding for global variable ‘marker2’

> calc_lod: no visible binding for global variable ‘log10detrss’

> Undefined global functions or variables:

>  log10detrss marker1 marker2

I believe that the note arises because of the nonstandard evaluation methods that the `dplyr` R package uses.

I then sought to fix the code so that, on running `devtools::check()`, I would not create this note.

For the reader to understand my goals with this function, I will take a moment to describe the inputs to `calc_lod`. I use two inputs, `data` and `n_mouse`. The first, `data`, is a tibble with three columns. The columns have the names: `marker1`, `marker2`, and `log10detrss`. Each row represents a unique (ordered) pairing of markers and contains, in the third column, the logarithm of the determinant of the residual sum of squares matrix. 

The second input, `n_mouse`, is needed for the calculation of LOD scores, as the code demonstrates. It corresponds to the number of mice used in the analysis.

## Programming with `dplyr` R package

The [`dplyr` programming vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/programming.html) is where I started in my efforts to understand `dplyr` evaluation methods.

## Finding the answer in `dplyr` programming vignette

Near the top of the programming vignette, just above the heading "Different Expressions", the author writes:

> If this function is in a package, using .data also prevents R CMD check from giving a NOTE about undefined global variables (provided that you’ve also imported rlang::.data with @importFrom rlang .data).

After a google search, I realized that `rlang` is an R package that I've never before used. However, I still wasn't sure exactly how to accomplish what the vignette says is a solution. 

I decided to rewrite my function by adding `.data$` to the variable names as needed. This led me to have a function that looks like this:

```{r, eval = FALSE}
calc_lod <- function(data, n_mouse){
  # define log10detrss0
  log10detrss0 <- dplyr::filter(data, .data$marker1 == .data$marker2)
  pre1 <- dplyr::group_by(data, .data$marker1)
  profile1 <- dplyr::summarise(pre1, profile = min(.data$log10detrss))
  pre2 <- dplyr::group_by(data, .data$marker2)
  profile2 <- dplyr::summarise(pre2, profile = min(.data$log10detrss))
  return(tibble::tibble(lod1 = - n_mouse * (profile1$profile - min(log10detrss0$log10detrss)) / 2,
                        lod2 = - n_mouse * (profile2$profile - min(log10detrss0$log10detrss))/ 2,
                        joint = - n_mouse * (log10detrss0$log10detrss - min(log10detrss0$log10detrss)) / 2))
}
```

In re-reading the sentence:

> If this function is in a package, using .data also prevents R CMD check from giving a NOTE about undefined global variables (provided that you’ve also imported rlang::.data with @importFrom rlang .data).

I guessed that I needed to add to my R script for calc_lod a line of notation for roxygen2. 

With this line, my calc_lod.R file - which contains only the function `calc_lod` - looks like this:

```{r, eval = FALSE}
#' Calculate LODs - two profile and one joint - for two traits
#'
#' @param data tibble containing log10determinants of RSS
#' @param n_mouse sample size
#' @export
#' @importFrom rlang .data

calc_lod <- function(data, n_mouse){
  # define log10detrss0
  log10detrss0 <- dplyr::filter(data, .data$marker1 == .data$marker2)
  pre1 <- dplyr::group_by(data, .data$marker1)
  profile1 <- dplyr::summarise(pre1, profile = min(.data$log10detrss))
  pre2 <- dplyr::group_by(data, .data$marker2)
  profile2 <- dplyr::summarise(pre2, profile = min(.data$log10detrss))
  return(tibble::tibble(lod1 = - n_mouse * (profile1$profile - min(log10detrss0$log10detrss)) / 2,
                        lod2 = - n_mouse * (profile2$profile - min(log10detrss0$log10detrss))/ 2,
                        joint = - n_mouse * (log10detrss0$log10detrss - min(log10detrss0$log10detrss)) / 2))
}
```

It's important to notice that I manually added the line 

```{r, eval = FALSE}
#' @importFrom rlang .data
```

so that roxygen2, the R package that assembles the NAMESPACE file, would know to import `.data` from R package `rlang`. 

Finally, I added `rlang` to the `pleiotropy` package dependencies by typing:

```{r, eval = FALSE}
devtools::use_package("rlang")
```

Now, when I run `devtools::check()` in the `pleiotropy` package directory, I see that there is no note produced:

> R CMD check results

> 0 errors | 0 warnings | 0 notes

